The Cost of Exclusion: Political Discrimination and Trust Deficits in the European Union

### 2025-07-14 article/main.tex
First Draft ready

While Europe has historically exhibited lower levels of affective and ideological polarization, recent studies suggest the rising presence of partisan hostility and negative feelings driven by factors such as ideological distance, perceived party threats, and regional tensions. Yet, evidence on longitudinal trends is still inconclusive. Scholars continue to emphasize the need for improved measurements and long-term data to determine whether polarization is indeed increasing across European democracies \parencite{wagner_affective_2024}.

@@

\section{Literature Review}

@@

\medskip
\justifying\footnotesize 
\textit{Note:} The figure shows hypothetical sensitivity contours for the point estimates of the treatment effect of political discrimination and harassment on trust in political institutions. The contours represent the hypothetical residual share of variation of the treatment that an unobserved confounding explains (x-axis) and the the hypothetical partial $R^2$ of an unobserved confounding with the outcome (y-axis). The red line represents the level at which the unobserved confounding would nullify our results. The magnitude of the coefficient of interest (ATT) is shown within parenthesis.

>>>>

While Europe has historically exhibited lower levels of affective and ideological polarization, recent studies suggest the rising presence of partisan hostility and negative feelings driven by factors such as ideological distance, perceived party threats, and regional tensions \parencite{ reiljan_patterns_2024, torcal_ideological_2022, garzia_negative_2021}. Yet, evidence on longitudinal trends is still inconclusive. Scholars continue to emphasize the need for improved measurements and long-term data to determine whether polarization is indeed increasing across European democracies \parencite{wagner_affective_2024}.

@@

\section{Conceptual Framework and Research Expectations}

@@

\medskip
\justifying\footnotesize 
\textit{Note:} The figure shows hypothetical sensitivity contours for the point estimates of the treatment effect of political discrimination and harassment on trust in political institutions. The contours represent the hypothetical residual share of variation of the treatment that an unobserved confounding explains (x-axis) and the hypothetical partial $R^2$ of an unobserved confounding with the outcome (y-axis). The red line represents the level at which the unobserved confounding would nullify our results. The magnitude of the coefficient of interest (ATT) is shown within parenthesis.
